\section{The atlas pipeline}\label{sec:method}

The atlas is produced by a single deterministic pipeline that takes a length bound
and returns, for every handshake within that bound, a vector of invariants drawn from
three views of the recognition prefix. No step uses randomness, and every numeric
column is reproducible to the last digit from the prefix alone.

\subsection{Enumeration}

For a fixed length $k$ the prefix alphabet is $\{C,D\}$, so the stratum of length-$k$
handshakes has $2^{k}$ members. We enumerate each stratum exhaustively in big-endian
order, reading $C$ as the low symbol and $D$ as the high symbol within a fixed-width
binary counter, and concatenate strata by increasing $k$. The order is total and
canonical, so two runs of the pipeline visit handshakes in the same sequence and assign
the same index to each. Structural invariants are computed for every handshake up to
$k\le 13$, which is $\sum_{k=1}^{13} 2^{k}=16{,}382$ handshakes. The evolutionary view,
which requires a pairwise match per handshake, is computed up to $k\le 8$, which is
$\sum_{k=1}^{8} 2^{k}=510$ handshakes. Both bounds are exhaustive within their stratum:
no sampling, no pruning.

\subsection{The joint invariant battery}

The invariants are grouped into three views, matching the three ways a recognition
prefix can be read: as a string, as a machine, and as a player.

\paragraph{String view.}
The prefix $p$ is mapped to its $\pm1$ image and profiled as a finite binary sequence.
We record the aperiodic autocorrelation sidelobes $c_v=\sum_j a_j a_{j+v}$, the peak
sidelobe $\max_v |c_v|$, and the merit factor $N^2/(2\sum_v c_v^2)$; a handshake is a
Barker code when its peak sidelobe is at most $1$ \citep{barker1953,turyn1961}. From the
Guibas--Odlyzko self-overlap (correlation) polynomial we read whether $p$ has a proper
border, that is a proper prefix equal to a suffix; a prefix with no proper border is
unbordered (bifix-free) \citep{guibas1981}. Compressibility is measured three ways:
LZ76 parsing complexity \citep{lempel1976}, linear complexity via the Berlekamp--Massey
recurrence \citep{massey1969}, and factor complexity. These columns are exact integer or
rational functions of $p$.

\paragraph{Automaton view.}
The gated strategy $H_p$ is realised as a Moore machine. The construction has a spine of
$k$ states that play out $p$ and check the opponent against it round by round, a
post-recognition regime $\rho$, and a defect-forever sink, giving an unminimised machine
of $2k+2$ states. We then minimise by partition refinement: states are split until no
block contains two states with different outputs or with a transition leaving the block
on some symbol, which is exactly the Myhill--Nerode quotient
\citep{nerode1958,hopcroft1971}. The resulting state count $m(p)$ is the
Rubinstein strategy-complexity measure of $H_p$ \citep{rubinstein1986}. Because the
refinement is canonical, $m(p)$ is a deterministic function of $p$ and $\rho$.

\paragraph{Evolutionary view.}
For the evolutionary columns each handshake is played in a deterministic pairwise match
against a fixed reference set under the payoffs $R=3,\,S=0,\,T=5,\,P=1$. Since both
players are deterministic, the round-by-round play and therefore the exact mean
per-round payoff over $n$ rounds ($n=200$ by default) are fixed, with no Monte Carlo
estimate. These mean payoffs are the entries of a constant-selection fitness comparison,
which we feed to the closed-form fixation probability of a single mutant in a
finite Moran population \citep{moran1958,nowak2004}, with default population $N=20$ and
selection intensity $w=0.1$; the neutral-drift baseline is $1/N$. Two tests use this
machinery. The \emph{invade-AllD} test computes the fixation probability of $H_p$ under
the \textsc{coop} regime invading a population of unconditional defectors, isolating the
cost of carrying a recognition prefix. The \emph{mimic-resistance} test computes the
fixation probability of a mimic of $p$ (which plays $p$ then defects forever) against a
resident population of $H_p$, run once per post-recognition regime
$\rho\in\{\textsc{coop},\textsc{grim},\textsc{tft}\}$, to ask whether the regime punishes
a forged signal.

\subsection{Determinism and reproducibility}

Every stage is a deterministic function of the prefix and a small fixed parameter set.
The enumeration order is canonical, the minimisation is canonical, the match play is
deterministic, and the fixation probabilities are evaluated in closed form rather than
simulated. Re-running the pipeline reproduces every number exactly; there is no random
seed to fix.

\begin{algorithm}[t]
\caption{The atlas pipeline for length bound $K$}
\label{alg:pipeline}
\begin{algorithmic}[1]
\For{$k = 1$ to $K$}
  \For{each $p \in \{C,D\}^k$ in big-endian order}
    \State \textbf{String:} compute autocorrelation sidelobes, peak sidelobe, merit
      factor, Barker flag; self-overlap and unbordered flag; LZ76, linear, and factor
      complexity.
    \State \textbf{Automaton:} build the $2k+2$-state Moore machine for $H_p$;
      partition-refine to obtain $m(p)$.
    \If{$k \le 8$}
      \State \textbf{Evolutionary:} run the deterministic match for exact mean payoffs;
        evaluate the closed-form Moran fixation for invade-AllD and for
        mimic-resistance under each $\rho$.
    \EndIf
    \State emit the invariant row for $p$.
  \EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\subsection{Scope of the distance column}

One natural invariant is deliberately absent at the per-handshake level. Within a full
length-$k$ stratum every prefix has a neighbour at Hamming distance $1$ (flip any single
bit), so the minimum distance from a handshake to the rest of its stratum is trivially
$1$ and carries no information. We therefore treat Hamming distance as a set-level
quantity, useful for separating families or code sets, and not as a per-handshake
column in the atlas.

\section{A taxonomy of handshake families}\label{sec:families}

The invariants induce a coarse classification of handshakes into six families. Each
family is a predicate over the invariant vector, and a handshake is assigned to the
first family whose predicate it satisfies, in the order Constant, Alternating, Barker,
Unbordered, Self-similar, Generic. The first-match rule makes the assignment a total
function, with Generic as the catch-all for prefixes that no earlier predicate selects.
\Cref{tab:families} states the predicates.

\begin{table}[t]
\caption{The six handshake families, defined as priority predicates over the invariants.
A handshake is assigned to the first family whose predicate it satisfies, top to bottom.}
\label{tab:families}
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Family & Defining predicate & Characterisation \\
\midrule
Constant     & $p \in \{C^k, D^k\}$                          & one repeated symbol \\
Alternating  & $p = (CD)^{\ast}$ or $(DC)^{\ast}$ truncated  & period-two pattern \\
Barker       & peak autocorrelation sidelobe $\le 1$         & flat-sidelobe code \\
Unbordered   & no proper border (bifix-free)                 & no prefix-suffix overlap \\
Self-similar & low LZ76 / linear complexity, not above       & compressible, repetitive \\
Generic      & none of the above                             & incompressible bulk \\
\bottomrule
\end{tabular}
\end{table}

\subsection{The taxonomy recovers the geometry}

The families are rule-based, so it is fair to ask whether they correspond to any
structure in the invariants themselves or merely reflect the thresholds chosen.
Clustering the standardised invariant vectors directly answers this.

\begin{observation}[$k$-means recovers the rule families]
\label{obs:kmeans}
For $k\le 8$, $k$-means over the standardised joint-invariant vectors recovers the
rule-based families as distinct clusters: one cluster is $100\%$ Constant, one is $85\%$
Barker, two are $88$--$100\%$ Unbordered, and one is $76\%$ Self-similar. The
unsupervised geometry therefore aligns with the priority predicates, rather than the
predicates imposing arbitrary cuts on an otherwise structureless space.
\end{observation}

\subsection{A thin structured shell}

The structured families do not grow with the space. There are always exactly two
Constant handshakes, $C^k$ and $D^k$, at every length, and for $k\ge 3$ there are always
exactly two Alternating handshakes. The stratum total, by contrast, doubles at each
length. The share of these structured families therefore tends to zero: the recognisable,
compressible prefixes form a thin shell inside an exponentially growing bulk of Generic
handshakes. The same picture holds for the unbordered fraction, which decreases along the
length axis and converges from above toward the known bifix-free density
$\approx 0.2678$ \citep{oeisA003000}; the special structure is a vanishing minority of
the recognition space.
